---
title: "ESM204 HW2"
author: "Sarah Lam"
date: "4/20/2022"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(effsize)
library(kableExtra)
library(broom)
```

```{r}
# Read in the data
water_curves <- read_csv("Water_Districts.csv")
```
 
### ESM 204, Homework #2 Simulating a Water Market
Agricultural water is managed by irrigation districts and is increasingly scarce. To conserve water and reduce costs, four irrigation districts are considering developing a water market, in which water rights would be traded across the districts. In this homework you will conduct data analysis, simulations, and policy analysis to help inform the development of the water market. The dataset for this assignment is available in the file “Water_Districts.csv”, available on Gauchospace. The dataset contains estimates of the marginal cost of reducing water use in each of the four irrigation districts (Kern, Mojave, Antelope, and Ventura) for different levels of reduction (in Acre Feet). A few notes:
- Current (i.e. baseline) water use in these districts1 is: Kern = 150 AF, Mojave = 140 AF, Antelope = 220 AF, and Ventura = 245 AF. 
- Prices are in $/Acre Foot and quantities are in Acre Feet (AF). Please write a concise report that addresses all of the points below.

1. For each irrigation district, plot the marginal abatement cost data (from the data provided) and estimate a linear regression model with zero intercept. These estimated linear regressions will be your estimates of the marginal abatement cost curve (one for each irrigation district), which you will use for the remainder of this analysis.
.
```{r}
#individual plots
ggplot(data = water_curves, aes(x = Reduction, y = MC_Kern)) +
  geom_point(alpha = 0.5)
ggplot(data = water_curves, aes(x = Reduction, y = MC_Mojave)) +
  geom_point(alpha = 0.5)
ggplot(data = water_curves, aes(x = Reduction, y = MC_Antelope)) +
  geom_point(alpha = 0.5)
ggplot(data = water_curves, aes(x = Reduction, y = MC_Ventura)) +
  geom_point(alpha = 0.5)
ggplot(data = water_curves, aes(x = Reduction, y = MC_aggregate)) +
  geom_point(alpha = 0.5)
#combined plot 
ggplot() +
  geom_point(data = water_curves, aes(x = Reduction, y = MC_Kern, color = "Kern")) +
  geom_point(data = water_curves, aes(x = Reduction, y = MC_Mojave, color = "Mojave")) +
  geom_point(data = water_curves, aes(x = Reduction, y = MC_Antelope, color = "Antelope")) +
  geom_point(data = water_curves, aes(x = Reduction, y = MC_Ventura, color = "Ventura")) +
  labs(y = "$ / Acre Foot", 
       x = "Water Reduction (AF)", 
       title = "Marginal Cost of Water Reduction")
```

```{r}
###linear regression and demand for each district 
## Kern
mod_kern <- lm(MC_Kern ~ 0 + Reduction, data = water_curves) #regression line fit
summary(mod_kern) # equation from this is P = 2.29A
kern_baseline <- 150
kern_data <- tibble(Reduction = 0:150) # creates a vector of equally spaced numbers
predicted_MC <- predict(mod_kern, newdata = kern_data)
kern_data <- kern_data %>% 
  mutate(predicted_MC = predicted_MC)
# Plotting with predicted line
ggplot(data = water_curves, aes(x = Reduction, y = MC_Kern)) +
  geom_point(alpha = 0.5) +
  geom_line(data = kern_data, aes(x=Reduction, y=predicted_MC))
#Define demand by subtracting the reduction from the baseline - E0-E
kern_data <- kern_data %>% 
  mutate(water_demand = kern_baseline - Reduction)
# Plot the demand - P = 343.5 - 2.29Q
ggplot(kern_data, aes(x = water_demand, y = predicted_MC)) +
  geom_line() + 
  labs(x = 'water demand (AF)', y = '$/AF')

##Mojave
mod_mojave <- lm(MC_Mojave ~ 0 + Reduction, data = water_curves) #equation - P = 3.8A
mojave_baseline <- 140
mojave_data <- tibble(Reduction = 0:140) # creates a vector of equally spaced numbers
predicted_MC <- predict(mod_mojave, newdata = mojave_data)
mojave_data <- mojave_data %>% 
  mutate(predicted_MC = predicted_MC)
# Plotting with predicted line
ggplot(data = water_curves, aes(x = Reduction, y = MC_Mojave)) +
  geom_point(alpha = 0.5) +
  geom_line(data = mojave_data, aes(x=Reduction, y=predicted_MC))
mojave_data <- mojave_data %>% 
  mutate(water_demand = mojave_baseline - Reduction) # E0 - E
# Plot the demand - P = 532 - 3.8Q
ggplot(mojave_data, aes(x = water_demand, y = predicted_MC)) +
  geom_line() + 
  labs(x = 'water demand', y = '$/AF')

##Antelope
mod_antelope <- lm(MC_Antelope ~ 0 + Reduction, data = water_curves) #equation - P = 2.86A
antelope_baseline <- 220
antelope_data <- tibble(Reduction = 0:220) # creates a vector of equally spaced numbers
predicted_MC <- predict(mod_antelope, newdata = antelope_data)
antelope_data <- antelope_data %>% 
  mutate(predicted_MC = predicted_MC)
# Plotting with predicted line
ggplot(data = water_curves, aes(x = Reduction, y = MC_Antelope)) +
  geom_point(alpha = 0.5) +
  geom_line(data = antelope_data, aes(x=Reduction, y=predicted_MC))
antelope_data <- antelope_data %>%
  mutate(water_demand = antelope_baseline - Reduction) # E0 - E
# Plot the demand - P = 629.2 - 2.86Q
ggplot(antelope_data, aes(x = water_demand, y = predicted_MC)) +
  geom_line() + 
  labs(x = 'water demand (AF)', y = '$/AF')

##Ventura 
mod_ventura <- lm(MC_Ventura ~ 0 + Reduction, data = water_curves) #equation - P = 1.78A
ventura_baseline <- 245
ventura_data <- tibble(Reduction = 0:245) # creates a vector of equally spaced numbers
predicted_MC <- predict(mod_ventura, newdata = ventura_data)
ventura_data <- ventura_data %>% 
  mutate(predicted_MC = predicted_MC)
# Plotting with predicted line
ggplot(data = water_curves, aes(x = Reduction, y = MC_Ventura)) +
  geom_point(alpha = 0.5) +
  geom_line(data = ventura_data, aes(x=Reduction, y=predicted_MC))
ventura_data <- ventura_data %>% 
  mutate(water_demand = ventura_baseline - Reduction) # E0-E
# Plot the demand - P = 436.1 - 1.78Q
ggplot(ventura_data, aes(x = water_demand, y = predicted_MC)) +
  geom_line() + 
  labs(x = 'water demand (AF)', y = '$/AF')
```

```{r, eval=FALSE}
#regression lines and marginal cost plots 
ggplot() +
  geom_point(data = water_curves, aes(x = Reduction, y = MC_Kern, color = "Kern")) +
  geom_line(data = kern_data, aes(x=Reduction, y=predicted_MC, color = "Kern")) +
  geom_point(data = water_curves, aes(x = Reduction, y = MC_Mojave, color = "Mojave")) +
  geom_line(data = mojave_data, aes(x=Reduction, y=predicted_MC, color = "Mojave")) +
  geom_point(data = water_curves, aes(x = Reduction, y = MC_Antelope, color = "Antelope")) +
  geom_line(data = antelope_data, aes(x=Reduction, y=predicted_MC, color = "Antelope")) +
  geom_point(data = water_curves, aes(x = Reduction, y = MC_Ventura, color = "Ventura")) +
  geom_line(data = ventura_data, aes(x=Reduction, y=predicted_MC, color = "Ventura")) +
  labs(y = "$ / Acre Foot", 
       x = "Water Reduction (AF)", 
       title = "Marginal Cost of Water Reduction")
```

2. Using your estimated marginal abatement cost curves, derive each district’s demand curve for water. In other words, how much would each district be willing to pay for the right to use the first AF of water, second AF, etc.? Plot each district’s demand curve on a single graph. Which sector is willing to pay the most for the first AF of water?

```{r}
#### We want the data for all districts to be in one plot now
joined_demand <- bind_rows('Mojave' = mojave_data, 
                           'Kern' = kern_data, 
                           'Antelope' = antelope_data, 
                           'Ventura' = ventura_data, .id = 'Districts')
### plot the demand curves
ggplot(data = joined_demand, aes(x = water_demand, y = predicted_MC, color = Districts)) + 
  geom_line() +
  labs(title = "Predicted Marginal Demand Curves for Each District", 
       x = "Water Demand (AF)", 
       y = "$/AF")
```

3. Here you will analyze three alternative policies for reducing water use among these irrigation districts. In total, these irrigation districts will need to reduce water consumption from the current 755 AF down to 500 AF. For each intervention listed below, perform the following calculations: (1) calculate the stringency of the policy (defined below for each intervention) required to achieve the 500 AF target, (2) calculate the total cost of meeting the target using that approach, (3) calculate the cost to each district, and (4) calculate the tax revenue generated.

a. Cap without trade. Reduce each district’s water use by the same fraction (e.g., 1/3 of current baseline use), so the 500 AF target is met. Trade is not allowed. Stringency is defined as the magnitude of the fraction.
1) Stringency of the policy 
755-500 = 255.
p(150) + p(140)+ p(220) + p(245) = 255
p = 0.338 
2)Total cost of meeting the target: $2938.55 + $4278.78 + $7900.53 + $6089.61 = $21207.47
3)Cost to each district:
```{r}
# A function to compute the cost for a districts abatement given the abatement number
quantity_abate = 50
lm_mod = mod_kern
# defining the function
calc_cost <- function(lm_mod, quantity_abate) {
  estimated_cost = lm_mod$coefficients * quantity_abate
  total_cost = 0.5 * quantity_abate * estimated_cost
  return(total_cost[[1]])
}
# Q3a, kern abated 0.338 * 150 = 50.7 units
kern_cost <- calc_cost(mod_kern, quantity_abate = 50.7) 
mojave_cost <- calc_cost(mod_mojave, quantity_abate = 47.43) 
antelope_cost <- calc_cost(mod_antelope, quantity_abate = 74.36) 
ventura_cost <- calc_cost(mod_ventura, quantity_abate = 82.81) #these calculations are the same thing as if we calculated the area under the curve using the area of a triangle

district_costs <- tibble(kern_cost, mojave_cost, antelope_cost, ventura_cost) #put costs into nice table
district_costs %>% 
  kableExtra::kable(col.names = c("Kern",
                                  "Mojave", "Antelope", "Ventura")) %>% 
   kable_styling(full_width = FALSE)
```

4) There will be no tax revenue generated because there is no tax here. 

b. Tax on water use. A single tax is implemented on all water use in each of the four districts, such that water use is reduced to the 500 AF target. Stringency is defined as the magnitude of the tax.
We started by graphing the aggregate of the marginal cost curves for each water district. We did this by finding the slope of the aggregate line. 
P = 1.78A
**** Add in other formulas later 
MCagg = A(1/2.29 +1/3.9 +1/2.86 + 1/1.78)
A = .621, this is our slope of the aggregate line which we can graph using abline()

1) Stringency of the policy
The tax should be set at $158.36 for a reduction of 255 AF. 
```{r, }
aggMC <- function(x){
  return(0.621*x)
}
aggMC(255)
```

2) calculate the total cost of meeting the target using that approach
The total cost of the policy is the area under the aggregate line up to the reduction amount. 
This is 255 * 158.355 * .05 = $2019.03

```{r}
255*158.355*.05
```

3) calculate the cost to each district
The cost to each district is the area under the marginal cost curve for each district at the amount of the tax ($158.36). 
kern: 158.355/2.29 * 158.36 * 0.05 = 547.5349
mojave: 158.355/3.8 * 158.36 * 0.05 * 0.05 = 329.9618
antelope: 158.355/2.86 * 158.36 * 0.05 * 0.05 = 438.4108
ventura: 158.355/1.78 * 158.36 * 0.05 = 704.4128

```{r}
#kern 
158.355/2.29 * 158.36 * 0.05
#mojave
158.355/3.8 * 158.36 * 0.05
#antelope
158.355/2.86 * 158.36 * 0.05
#ventura
158.355/1.78 * 158.36 * 0.05

#Summing Q's
547.5349 +  329.9618 +  438.4108 + 704.4129 
2020.32
```

4) calculate the tax revenue generated.
Whole rectangle is tax revenue 

```{r}
ggplot() +
  geom_line(data = kern_data, aes(x=Reduction, y=predicted_MC, color = "Kern")) +
  geom_line(data = mojave_data, aes(x=Reduction, y=predicted_MC, color = "Mojave")) +
  geom_line(data = antelope_data, aes(x=Reduction, y=predicted_MC, color = "Antelope")) +
  geom_line(data = ventura_data, aes(x=Reduction, y=predicted_MC, color = "Ventura")) +
  geom_abline(aes(intercept = 0, slope = 0.621, colour = "Aggregate")) +
  scale_color_brewer(palette = "Dark2", 
                     breaks = c("Kern", "Mojave", "Antelope", "Ventura", "Aggregate")) +
  xlim(0,300) +
  ylim(0,300) +
  labs(y = "$ / Acre Foot", 
       x = "Water Reduction (AF)", 
       title = "Marginal Cost of Water Reduction",
       col = "District") 

#Tax Revenue: 
255 * 158.36
```

c. Cap and trade. Cap water use as in part (a), but after those caps are set, allow trade across districts. How much water is used by each district after trade? Stringency is the same as in part (a).


4. A severe drought hits California, and it is necessary to reduce water use from 500 AF down to 300 AF. Your job is to estimate the cost of the drought (i.e., the cost of reducing water use from 500 AF to 300 AF) to each irrigation district under each policy. Considering the entire jump from status quo (755 AF) to the drought (300 AF), which policy is preferred by each irrigation district? How does your answer depend on how rights are allocated to each district (in policies (a) and (c))2?












