---
title: "ESM204 HW2"
author: "Sarah Lam"
date: "4/20/2022"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(effsize)
library(kableExtra)
library(broom)
```

```{r}
# Read in the data
water_curves <- read_csv("Water_Districts.csv")
```
 
### ESM 204, Homework #2 Simulating a Water Market
Agricultural water is managed by irrigation districts and is increasingly scarce. To conserve water and reduce costs, four irrigation districts are considering developing a water market, in which water rights would be traded across the districts. In this homework you will conduct data analysis, simulations, and policy analysis to help inform the development of the water market. The dataset for this assignment is available in the file “Water_Districts.csv”, available on Gauchospace. The dataset contains estimates of the marginal cost of reducing water use in each of the four irrigation districts (Kern, Mojave, Antelope, and Ventura) for different levels of reduction (in Acre Feet). A few notes:
- Current (i.e. baseline) water use in these districts1 is: Kern = 150 AF, Mojave = 140 AF, Antelope = 220 AF, and Ventura = 245 AF. 
- Prices are in $/Acre Foot and quantities are in Acre Feet (AF). Please write a concise report that addresses all of the points below.

1. For each irrigation district, plot the marginal abatement cost data (from the data provided) and estimate a linear regression model with zero intercept. These estimated linear regressions will be your estimates of the marginal abatement cost curve (one for each irrigation district), which you will use for the remainder of this analysis.
.
```{r}
#individual plots
ggplot(data = water_curves, aes(x = Reduction, y = MC_Kern)) +
  geom_point(alpha = 0.5)
ggplot(data = water_curves, aes(x = Reduction, y = MC_Mojave)) +
  geom_point(alpha = 0.5)
ggplot(data = water_curves, aes(x = Reduction, y = MC_Antelope)) +
  geom_point(alpha = 0.5)
ggplot(data = water_curves, aes(x = Reduction, y = MC_Ventura)) +
  geom_point(alpha = 0.5)
ggplot(data = water_curves, aes(x = Reduction, y = MC_aggregate)) +
  geom_point(alpha = 0.5)
#combined plot 
ggplot() +
  geom_point(data = water_curves, aes(x = Reduction, y = MC_Kern, color = "Kern")) +
  geom_point(data = water_curves, aes(x = Reduction, y = MC_Mojave, color = "Mojave")) +
  geom_point(data = water_curves, aes(x = Reduction, y = MC_Antelope, color = "Antelope")) +
  geom_point(data = water_curves, aes(x = Reduction, y = MC_Ventura, color = "Ventura")) +
  labs(y = "$ / Acre Foot", 
       x = "Water Reduction (AF)", 
       title = "Marginal Cost of Water Reduction")

## Kern
mod_kern <- lm(MC_Kern ~ 0 + Reduction, data = water_curves) #regression line fit
summary(mod_kern)
kern_baseline <- 150
kern_data <- tibble(Reduction = 0:150) # creates a vector of equally spaced numbers
predicted_MC <- predict(mod_kern, newdata = kern_data)
kern_data <- kern_data %>% 
  mutate(predicted_MC = predicted_MC)
# Plotting with predicted line
ggplot(data = water_curves, aes(x = Reduction, y = MC_Kern)) +
  geom_point(alpha = 0.5) +
  geom_line(data = kern_data, aes(x=Reduction, y=predicted_MC))
#Define demand or use as Baseline (no policy) - Abatement
##### We use the data we already have, and mutate
kern_data <- kern_data %>% 
  mutate(water_demand = kern_baseline - Reduction)
# Plot the demand
ggplot(kern_data, aes(x = water_demand, y = predicted_MC)) +
  geom_line() + 
  labs(x = 'water demand (AF)', y = '$/AF')

##Mojave
mod_mojave <- lm(MC_Mojave ~ 0 + Reduction, data = water_curves)
mojave_baseline <- 140
mojave_data <- tibble(Reduction = 0:140) # creates a vector of equally spaced numbers
predicted_MC <- predict(mod_mojave, newdata = mojave_data)
mojave_data <- mojave_data %>% 
  mutate(predicted_MC = predicted_MC)
# Plotting with predicted line
ggplot(data = water_curves, aes(x = Reduction, y = MC_Mojave)) +
  geom_point(alpha = 0.5) +
  geom_line(data = mojave_data, aes(x=Reduction, y=predicted_MC))
#Define demand or use as Baseline (no policy) - Abatement
##### We use the data we already have, and mutate
mojave_data <- mojave_data %>% 
  mutate(water_demand = mojave_baseline - Reduction)
# Plot the demand
ggplot(mojave_data, aes(x = water_demand, y = predicted_MC)) +
  geom_line() + 
  labs(x = 'water demand', y = '$/AF')

##Antelope
mod_antelope <- lm(MC_Antelope ~ 0 + Reduction, data = water_curves)
antelope_baseline <- 220
antelope_data <- tibble(Reduction = 0:220) # creates a vector of equally spaced numbers
predicted_MC <- predict(mod_antelope, newdata = antelope_data)
antelope_data <- antelope_data %>% 
  mutate(predicted_MC = predicted_MC)
# Plotting with predicted line
ggplot(data = water_curves, aes(x = Reduction, y = MC_Antelope)) +
  geom_point(alpha = 0.5) +
  geom_line(data = antelope_data, aes(x=Reduction, y=predicted_MC))
#Define demand or use as Baseline (no policy) - Abatement
##### We use the data we already have, and mutate
antelope_data <- antelope_data %>% 
  mutate(water_demand = antelope_baseline - Reduction)
# Plot the demand
ggplot(antelope_data, aes(x = water_demand, y = predicted_MC)) +
  geom_line() + 
  labs(x = 'water demand (AF)', y = '$/AF')

##Ventrua 
mod_ventura <- lm(MC_Ventura ~ 0 + Reduction, data = water_curves)
ventura_baseline <- 245
ventura_data <- tibble(Reduction = 0:245) # creates a vector of equally spaced numbers
predicted_MC <- predict(mod_ventura, newdata = ventura_data)
ventura_data <- ventura_data %>% 
  mutate(predicted_MC = predicted_MC)
# Plotting with predicted line
ggplot(data = water_curves, aes(x = Reduction, y = MC_Ventura)) +
  geom_point(alpha = 0.5) +
  geom_line(data = ventura_data, aes(x=Reduction, y=predicted_MC))
#Define demand or use as Baseline (no policy) - Abatement
##### We use the data we already have, and mutate
ventura_data <- ventura_data %>% 
  mutate(water_demand = ventura_baseline - Reduction)
# Plot the demand
ggplot(ventura_data, aes(x = water_demand, y = predicted_MC)) +
  geom_line() + 
  labs(x = 'water demand (AF)', y = '$/AF')
```
```{r}
#regression lines and marginal cost plots 
ggplot() +
  geom_point(data = water_curves, aes(x = Reduction, y = MC_Kern, color = "Kern")) +
  geom_line(data = kern_data, aes(x=Reduction, y=predicted_MC, color = "Kern")) +
  geom_point(data = water_curves, aes(x = Reduction, y = MC_Mojave, color = "Mojave")) +
  geom_line(data = mojave_data, aes(x=Reduction, y=predicted_MC, color = "Mojave")) +
  geom_point(data = water_curves, aes(x = Reduction, y = MC_Antelope, color = "Antelope")) +
  geom_line(data = antelope_data, aes(x=Reduction, y=predicted_MC, color = "Antelope")) +
  geom_point(data = water_curves, aes(x = Reduction, y = MC_Ventura, color = "Ventura")) +
  geom_line(data = ventura_data, aes(x=Reduction, y=predicted_MC, color = "Ventura")) +
  labs(y = "$ / Acre Foot", 
       x = "Water Reduction (AF)", 
       title = "Marginal Cost of Water Reduction")
```

2. Using your estimated marginal abatement cost curves, derive each district’s demand curve for water. In other words, how much would each district be willing to pay for the right to use the first AF of water, second AF, etc.? Plot each district’s demand curve on a single graph. Which sector is willing to pay the most for the first AF of water?

```{r}
#### We want the data for all districts to be in one plot now
joined_demand <- bind_rows('Mojave' = mojave_data, 
                           'Kern' = kern_data, 
                           'Antelope' = antelope_data, 
                           'Ventura' = ventura_data, 
                           'Aggregate' = agg_data, .id = 'Districts')
### plot the demand curves
ggplot(data = joined_demand, aes(x = water_demand, y = predicted_MC, color = Districts)) + 
  geom_line() +
  labs(title = "Predicted Marginal Demand Curves for Each District", 
       x = "Water Demand (AF)", 
       y = "$/AF")
```

3. Here you will analyze three alternative policies for reducing water use among these irrigation districts. In total, these irrigation districts will need to reduce water consumption from the current 755 AF down to 500 AF. For each intervention listed below, perform the following calculations: (1) calculate the stringency of the policy (defined below for each intervention) required to achieve the 500 AF target, (2) calculate the total cost of meeting the target using that approach, (3) calculate the cost to each district, and (4) calculate the tax revenue generated.
a. Cap without trade. Reduce each district’s water use by the same fraction (e.g., 1/3 of current baseline use), so the 500 AF target is met. Trade is not allowed. Stringency is defined as the magnitude of the fraction.
b. Tax on water use. A single tax is implemented on all water use in each of the four districts, such that water use is reduced to the 500 AF target. Stringency is defined as the magnitude of the tax.
c. Cap and trade. Cap water use as in part (a), but after those caps are set, allow trade across districts. How much water is used by each district after trade? Stringency is the same as in part (a).
#### 3A, B, C

#All hinge on the demand curve. Once you have the final colorful plot you can tackle 3 A, B, and C.

### 3A
#We want a single number. 

```{r}
baseline = c(150, 140, 220, 245)
# A single fraction multiplied by these numbers that sum to 755-500 = 255. We want to deduct by a total of 255. 
# Use wolfram alpha to solve: p * 150 + p * 140 + p * 220 + p * 245 = 255
# p = 0.338


# A function to compute the cost for a districts abatement given the abatement number

quantity_abate = 50
lm_mod = mod_kern
# defining the function
calc_cost <- function(lm_mod, quantity_abate) {
  estimated_cost = lm_mod$coefficients * quantity_abate
  total_cost = 0.5 * quantity_abate * estimated_cost
  return(total_cost[[1]])
}

# Now with the above code we can answer questions like: What is the total cost if Kern were able to abate 117 units?
calc_cost(mod_kern, quantity_abate = 117)
# we can fill in different numbers for the 117, and get the total coast answer

# Q3a, kern abated 0.338 * 150 = 50.7 units
calc_cost(mod_kern, quantity_abate = 50.7) 
calc_cost(mod_mojave, quantity_abate = 47.43) 
calc_cost(mod_antelope, quantity_abate = 74.36) 
calc_cost(mod_ventura, quantity_abate = 82.81) 
# these calculations are the same thing as if we calculated the area under the curve up to 50.7 by using the area of a triangle. The answer we got here was 2938.55
```

4. A severe drought hits California, and it is necessary to reduce water use from 500 AF down to 300 AF. Your job is to estimate the cost of the drought (i.e., the cost of reducing water use from 500 AF to 300 AF) to each irrigation district under each policy. Considering the entire jump from status quo (755 AF) to the drought (300 AF), which policy is preferred by each irrigation district? How does your answer depend on how rights are allocated to each district (in policies (a) and (c))2?












